A contrast of meta and metafor packages for meta-analyses in R

### 2020-03-28 ms/2020/contrastRmd.Rmd
added keywords
edit and then add stata

@@

### Introduction
Meta-analyses are common and powerful synthesis tools in science.  Typically in the natural sciences, meta-analyses are used as a mechanism to describe and aggregate quantitative evidence from a set of peer-reviewed, primary research publications [@RN3629, @RN4850].  A meta-analysis in the natural sciences, i.e. outside the health sciences, is comprised of a formalized systematic review and analysis of effect sizes and is termed a meta-analysis when statistics examining intervention efficacy are included [@RN3216].  In other fields, the terms systematic review and meta-analysis are used more interchangeable, and meta-statistics are often done on compiled randomized controlled trials or other relatively large datasets in addition to sets of data derived from peer-reviewed publications.  The derived data in the natural sciences (as the intended primary audience here) routinely extracts only the mean values or single point measures from publications [@RN4861].  The synthesis statistics were commonly done using MetaWin [@RN6194] or other GUI-based applications for a number of years.  More recently however, statistics in the fields of ecology and evolution for instance have increasingly moved to the programming language R [@RN6098], and synthesis statistics are no exception.  At least two R packages have risen to prominence for general meta-analytical statistics in the natural sciences - namely meta [@RN6176] and metafor [@RN6175].  Given that meta-analyses are also increasingly published in these same fields [@RN2189, @RN3629], a brief comment on the ecosystem of analytical choices that R provides is beneficial and timely.  We need synthesis to inform evidence-based decisioning, and meta-analyses can be the primary tool if aggregated primary datasets are unavailable.  Furthermore even with primary data in hand, data reduction to effect sizes within primary and synthesis studies is a mechanism to illustrate differences and strength of effects.  These approaches provide the capacity for higher-order analyses and reuse [@RN4835] suggesting that familiarity with effect sizes is both germane and practical.

>>>>

### Keywords
contrasts, computational tools, Meta, meta-analyses, Metafor, methods, R, review

@@

### Introduction
Meta-analyses are common and powerful synthesis tools in science.  Typically in the natural sciences, meta-analyses are used as a mechanism to describe and aggregate quantitative evidence from a set of peer-reviewed, primary research publications [@RN3629; @RN4850].  A meta-analysis in the natural sciences, i.e. outside the health sciences, is comprised of a formalized systematic review and analysis of effect sizes and is termed a meta-analysis when statistics examining intervention efficacy are included [@RN3216].  In other fields, the terms systematic review and meta-analysis are used more interchangeable, and meta-statistics are often done on compiled randomized controlled trials or other relatively large datasets in addition to sets of data derived from peer-reviewed publications.  The derived data in the natural sciences (as the intended primary audience here) routinely extracts only the mean values or single point measures from publications [@RN4861].  The synthesis statistics were commonly done using MetaWin [@RN6194] or other GUI-based applications for a number of years.  More recently however, statistics in the fields of ecology and evolution for instance have increasingly moved to the programming language R [@RN6098], and synthesis statistics are no exception.  At least two R packages have risen to prominence for general meta-analytical statistics in the natural sciences - namely meta [@RN6176] and metafor [@RN6175].  Given that meta-analyses are also increasingly published in these same fields [@RN2189; @RN3629], a brief comment on the ecosystem of analytical choices that R provides is beneficial and timely.  We need synthesis to inform evidence-based decisioning, and meta-analyses can be the primary tool if aggregated primary datasets are unavailable.  Furthermore even with primary data in hand, data reduction to effect sizes within primary and synthesis studies is a mechanism to illustrate differences and strength of effects.  These approaches provide the capacity for higher-order analyses and reuse [@RN4835] suggesting that familiarity with effect sizes is both germane and practical.
